---
title: "Thymus progenitor supplementary files - figure 2"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

This notebook constructs the various supplementary figure panels that correspond to main figure 2.

```{r}
suppressMessages({
library(ggplot2)
library(knitr)
library(kableExtra)
library(ggsci)
library(cowplot)
library(ggrepel)
library(scattermore)
library(SingleCellExperiment)
library(scran)
library(scater)
library(scuttle)
library(batchelor)
library(igraph)
library(miloR)
library(reshape2)
library(dplyr)
library(biomaRt)
library(BiocNeighbors)
library(ggalluvial)
})
```

# Single-cell QC

```{r "emptyDrops calling"}
edrops.df <- read.table("~/Dropbox/AgeingExperiment/Perinatal_manuscript/data/Perinatal_emptyDrops.tsv",
                        header=FALSE, sep="\t")
colnames(edrops.df) <- c("InFile", "Droplets")
edrops.df$SampName <- gsub(edrops.df$InFile, pattern="^(logs/)(Newborn)_(ZsG[a-z]+)_(\\S+Run)_([m|c]TEC)_(HTO-emptyDrops\\.err)",
                           replacement="\\3_\\4_\\5")
```

Plot the number of non-empty droplets called by `emptyDrops`.

```{r "emptyDrops plotting", fig.height=4.75, fig.width=6}
ggplot(edrops.df, aes(x=SampName, y=Droplets)) +
  geom_bar(stat='identity', colour='grey80', fill='grey80') +
  geom_text(aes(label=Droplets)) +
  theme_cowplot() +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
  labs(x="Sample ID", y="#non-empty droplets") +
  NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/EmptyDrops-calling-bar.png",
       height=4.75, width=6, dpi=300, bg='white')
```

```{r "high mitochondrial expression"}
# this uses an adaptive threshold based on the distribution of MT RNA expression - specifically cells with a significant departure from the 
# median absolute deviations after FDR adjustment
mito.df <- read.table("~/Dropbox/AgeingExperiment/Perinatal_manuscript/data/Perinatal_highMito.tsv",
                      sep="\t", header=TRUE)
mito.df$SampName <- gsub(mito.df$InFile, pattern="^(logs/)(Newborn)_(ZsG[a-z]+)_(\\S+Run)_([m|c]TEC)_(HTO-emptyDrops\\.err)",
                           replacement="\\3_\\4_\\5")
```

```{r "high mitochondrial QC plotting", fig.height=4.75, fig.width=6}
ggplot(mito.df, aes(x=SampName, y=high.Mito)) +
  geom_bar(stat='identity', colour='grey80', fill='grey80') +
  geom_text(aes(label=high.Mito)) +
  theme_cowplot() +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
  labs(x="Sample ID", y="#high mito. RNA cells") +
  NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/High_mitochondrial_RNA-bar.png",
       height=4.75, width=6, dpi=300, bg='white')
```

```{r "doublet calling from HTOs"}
hto.df <- read.table("~/Dropbox/AgeingExperiment/Perinatal_manuscript/data/HTO_class.tsv",
                     header=TRUE, sep="\t")
hto.df$SampName <- gsub(hto.df$ID, pattern="(Newborn)_(ZsG[a-z]+)_(\\S+Run)_([m|c]TEC)_(HTO_[ATCG]+-1)", replacement="\\2_\\3_\\4")
```


```{r "HTO doublet plotting", fig.height=5, fig.width=9}
set.seed(42)
# its easier to tabulate and summarise so the numbers can be added
hto.summ.df <- hto.df %>% group_by(SampName) %>% count(Class) %>% mutate(Prop.HTO=n/sum(n))

ggplot(hto.summ.df, aes(x=SampName, y=Prop.HTO, fill=Class)) +
  geom_bar(stat='identity', position="dodge") +
  geom_text(aes(y=Prop.HTO+0.1, label=round(n, 3), group=Class),
            position=position_dodge(width=1), size=4.5, angle=90) +
  theme_cowplot() +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1),
        ) +
  labs(x="Sample ID", y="Proportion in HTO class") +
  scale_fill_d3() +
  expand_limits(y=c(0, 1.15)) +
  NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/HTO_doublets-bar.png",
       height=5, width=9, dpi=300, bg='white')
```

# Comparison to Wells *et al.* data

```{r "Ensembl to gene symbol mapping"}
biomaRt.connection <- useMart("ensembl", "mmusculus_gene_ensembl")

gene.df <- getBM(attributes = c("ensembl_gene_id", "external_gene_name", "chromosome_name"),
                 mart=biomaRt.connection)
rownames(gene.df) <- gene.df$ensembl_gene_id
```

```{r "loading Wells et al data"}
all.controls <- read.csv("~/Dropbox/AgeingExperiment/Perinatal_manuscript/data/GSE137699_allSamples.csv.gz",
                         header=TRUE)

all.meta <- read.table("~/Dropbox/AgeingExperiment/Perinatal_manuscript/data/GSE137699_RAW/all_meta.csv",
                       header=TRUE, sep=",")
rownames(all.meta) <- all.meta$X
all.meta$stage[all.meta$stage %in% c("Aire_positive~")] <- "Aire_positive"

keep.cells <- intersect(rownames(all.meta), colnames(all.controls))

wells.sce <- SingleCellExperiment(assays=list(logcounts=as(as.matrix(all.controls[, keep.cells]), "dgCMatrix")), 
                                  colData=all.meta[keep.cells, ])

aire.iso.sce <- wells.sce[, colData(wells.sce)$exp %in% c("aireTrace", "isoControlBeg", "isoControlEnd")]

symb.to.ensembl <- gene.df[gene.df$external_gene_name %in% rownames(aire.iso.sce), c("external_gene_name", "ensembl_gene_id")]
symb.to.ensembl <- symb.to.ensembl[!duplicated(symb.to.ensembl$external_gene_name), ]
symb.to.ensembl <- symb.to.ensembl[!is.na(symb.to.ensembl$external_gene_name), ]
rownames(symb.to.ensembl) <- symb.to.ensembl$external_gene_name

aire.iso.sce <- aire.iso.sce[rownames(aire.iso.sce) %in% rownames(symb.to.ensembl), ]
aire.iso.sce <- aire.iso.sce[rownames(symb.to.ensembl) , ]
rownames(aire.iso.sce) <- symb.to.ensembl$ensembl_gene_id
```

I'll start with the Aire lineage traced and isotype control samples from
the antibody blockade experiments.

```{r "Aire-traced and isotype controls dimensionality reduction"}
set.seed(42)
aire.iso.hvgs <- modelGeneVar(aire.iso.sce)
aire.iso.hvgs$FDR[is.na(aire.iso.hvgs$FDR)] <- 1

reducedDim(aire.iso.sce, "PCA") <- multiBatchPCA(aire.iso.sce, subset.row=aire.iso.hvgs$FDR < 0.1, assay.type="logcounts",
                                                 batch=colData(aire.iso.sce)$exp, d=50, preserve.single=TRUE)[[1]]
aire.iso.sce <- runUMAP(aire.iso.sce, n_neighbors=21, dimred="PCA", n_dimred=30)
aire.iso.umap <- data.frame("CellID"=colnames(aire.iso.sce), "UMAP1"=reducedDim(aire.iso.sce, "UMAP")[, 1],
                            "UMAP2"=reducedDim(aire.iso.sce, "UMAP")[, 2])
aire.iso.umap <- merge(aire.iso.umap, all.meta, by.x="CellID", by.y="X")
```


```{r "UMAP Aire-traced and isotype controls", fig.height=3.75, fig.width=11}
by.stage <- ggplot(aire.iso.umap, aes(x=UMAP1, y=UMAP2, colour=stage)) +
  geom_scattermore(pointsize=1) +
  theme_cowplot() +
  scale_colour_d3() +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  NULL

by.batch <- ggplot(aire.iso.umap, aes(x=UMAP1, y=UMAP2, colour=exp)) +
  geom_scattermore(pointsize=1) +
  theme_cowplot() +
  scale_colour_npg() +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  NULL

plot_grid(by.stage, by.batch)
```

These UMAPs have similar structure to those shown in Wells *et al.*
figure 1, which is encouraging. The next step is to combine these with
our perinatal TEC data set and cross-compare the cell labels.

```{r "load perinatal data"}
# derive all of the relevant meta data
perinatal.milo <- readRDS("~/Dropbox/AgeingExperiment/Newborn/milo.dir/Perinatal_Milo.RDS")

colData(perinatal.milo)$SortType <- gsub(perinatal.milo$HTO, pattern="(ZsG[n|p])_(d[0-9]+)_(\\S+)_(R[0-4])", replacement="\\1")
colData(perinatal.milo)$TECSort <- gsub(perinatal.milo$HTO, pattern="(ZsG[n|p])_(d[0-9]+)_(\\S+)_(R[0-4])", replacement="\\3")
colData(perinatal.milo)$Day <- gsub(perinatal.milo$HTO, pattern="(ZsG[n|p])_(d[0-9]+)_(\\S+)_(R[0-4])", replacement="\\2")
colData(perinatal.milo)$Day[colData(perinatal.milo)$Day %in% c("d20")] <- "d2"
colData(perinatal.milo)$Replicate <- gsub(perinatal.milo$HTO, pattern="(ZsG[n|p])_(d[0-9]+)_(\\S+)_(R[0-4])", replacement="\\4")

perinatal.meta <- colData(perinatal.milo)
perinatal.meta$CellID <- colnames(perinatal.milo)
perinatal.umap <- data.frame("CellID"=colnames(perinatal.milo), "UMAP1"=reducedDim(perinatal.milo, "UMAP")[, 1],
                             "UMAP2"=reducedDim(perinatal.milo, "UMAP")[, 2])
perinatal.meta <- merge(perinatal.meta, perinatal.umap, by='CellID')

milo.meta <- read.table("~/Dropbox/AgeingExperiment/Perinatal_manuscript/results/Milo_res-ZsGCompare.txt",
                        sep="\t", header=TRUE)
milo.meta$NhoodIdx <- colnames(nhoods(perinatal.milo))
```

```{r, warning=FALSE, message=FALSE}
# first assign each cell to it's nhood(s)
cell.to.nhood <- do.call(rbind.data.frame, lapply(apply(nhoods(perinatal.milo), 1,
                                                        FUN=function(NX){
                                                            colnames(nhoods(perinatal.milo))[NX > 0]
                                                            }),
                                                  FUN=function(CN) paste(CN, collapse=",")))
colnames(cell.to.nhood) <- c("Nhoods")
cell.to.nhood$CellID <- colnames(perinatal.milo)

# add the nhood group(s)
cell.to.nhgroup <- data.frame("CellID"=cell.to.nhood$CellID, 
                              "NhoodGroup"=apply(cell.to.nhood, 1,
                                                 FUN=function(CX){
                                                     paste(unique(milo.meta[milo.meta$NhoodIdx %in% 
                                                                              unlist(strsplit(CX["Nhoods"], split=",", fixed=TRUE)),]$NhoodGroup),
                                                           collapse=",")      }),
                              "NhoodLabel"=apply(cell.to.nhood, 1,
                                                 FUN=function(CX){
                                                     paste(unique(milo.meta[milo.meta$NhoodIdx %in% 
                                                                              unlist(strsplit(CX["Nhoods"], split=",",
                                                                                              fixed=TRUE)),]$Nhood.TEC),
                                                           collapse=",")      }))
```

I've assigned each cell to it's nhoods, nhood groups and TEC labels.
Next I will resolve the cells which fall into multiple TEC groups. To do
this, I will find the cells assigned to multiple groups and look at
their nearest neighbours. If they are mostly assigned to a single
group/label I will assign the cell to this group/label.

```{r, warning=FALSE, message=FALSE}
cell.to.nhgroup$NLabels <- apply(cell.to.nhgroup, 1, FUN=function(LX){
    length(unlist(strsplit(LX["NhoodLabel"], split=",", fixed=TRUE)))
    })

non.unique.cells <- cell.to.nhgroup$CellID[cell.to.nhgroup$NLabels > 1]
non.unique.nn <- findKNN(reducedDim(perinatal.milo, "PCA"), k=perinatal.milo@.k,
                         get.distance=FALSE,
                         subset=which(colnames(perinatal.milo) %in% non.unique.cells))$index # should this be the same as the original graph/nhood definitions?
rownames(non.unique.nn) <- non.unique.cells

non.unique.labels <- list()
for(x in seq_along(non.unique.cells)){
    x.cell <- non.unique.cells[x]
    x.nn <- colnames(perinatal.milo)[non.unique.nn[x.cell, ]]
    x.tab <- table(cell.to.nhgroup[cell.to.nhgroup$CellID %in% x.nn & cell.to.nhgroup$NLabels == 1, ]$NhoodLabel)
    x.max <- names(x.tab)[which(x.tab == max(x.tab))]
    if(length(x.max) == 1){
        non.unique.labels[[x.cell]] <- x.max
    } else{
        non.unique.labels[[x.cell]] <- NA # if there is no unique value
    }
}
non.unique.df <- do.call(rbind.data.frame, non.unique.labels)
colnames(non.unique.df) <- ("UniqueLabel")
non.unique.df$CellID <- non.unique.cells

# less than 10% get an NA, that's not bad.
rownames(cell.to.nhgroup) <- cell.to.nhgroup$CellID
cell.to.nhgroup$UniqueLabel <- cell.to.nhgroup$NhoodLabel
cell.to.nhgroup$UniqueLabel[cell.to.nhgroup$NLabels > 1] <- NA
cell.to.nhgroup[non.unique.df$CellID, ]$UniqueLabel <- non.unique.df$UniqueLabel
cell.to.nhgroup <- cell.to.nhgroup[cell.to.nhgroup$UniqueLabel != "", ]

# merge with the meta-data
cell.to.nhgroup <- merge(cell.to.nhgroup, perinatal.meta[, c("SortType", "CellID", "Replicate")], by='CellID')
cell.to.nhgroup$ZsGLabel <- paste(cell.to.nhgroup$UniqueLabel, cell.to.nhgroup$SortType, sep="_")
```


```{r}
# annotate the single cells
rownames(cell.to.nhgroup) <- cell.to.nhgroup$CellID
colData(perinatal.milo)$Nhood.TEC <- cell.to.nhgroup[colnames(perinatal.milo), ]$UniqueLabel
colData(perinatal.milo)$Nhood.TEC <- factor(colData(perinatal.milo)$Nhood.TEC,
                                            levels=c("Immature cTEC (1)", "Cycling cTEC (2)", "cTEC late (3)", "cTEC early (4)",
                                                     "mTEC (5)", "Intertypical (6)", "Post-Aire mTEC (7)", "cTEC early (8)",
                                                     "ILC-like (9)", "Cycling mTEC (10)", "Sca1+ mTEC (11)", "cTEC late (12)",
                                                     "Intertypical cTEC-biased (13)", "ILC-like (14)", "Cycling Intertypical (15)",
                                                     "Tuft-like mTEC (16)", "ILC-like (17)", "Aire+ mTEC (18)"))

n.groups <- length(unique(milo.meta$NhoodGroup))
nhoodlabel.cols <- colorRampPalette(pal_d3("category20")(20))(n.groups)

names(nhoodlabel.cols) <- c("Immature cTEC (1)", "Cycling cTEC (2)", "cTEC late (3)", "cTEC early (4)", "mTEC (5)", "Intertypical (6)", "Post-Aire mTEC (7)", 
                            "cTEC early (8)", "ILC-like (9)", "Cycling mTEC (10)", "Sca1+ mTEC (11)", "cTEC late (12)",
                            "Intertypical cTEC-biased (13)", "ILC-like (14)", "Cycling Intertypical (15)", "Tuft-like mTEC (16)", "ILC-like (17)", "Aire+ mTEC (18)")

```

We will integrate the data sets using `reducedMNN`.

```{r "shared reduced dimesions on Perinatal and Wells et al data"}
set.seed(42)

keep.genes <- intersect(rownames(aire.iso.sce), rownames(perinatal.milo))
joint.sce <- SingleCellExperiment(assays=list(logcounts=do.call(cbind, list(logcounts(perinatal.milo[keep.genes, ]), logcounts(aire.iso.sce[keep.genes, ])))),
                                              colData=data.frame("Batch"=c(rep("Perinatal", ncol(perinatal.milo)), rep("Wells", ncol(aire.iso.sce)))))
joint.hvg <- modelGeneVar(joint.sce)
joint.hvg$FDR[is.na(joint.hvg$FDR)] <- 1
assay(joint.sce, "Cosine") <- cosineNorm(logcounts(joint.sce), mode="matrix")
reducedDim(joint.sce, "Cosine.PCA") <- multiBatchPCA(joint.sce, subset.row=joint.hvg$FDR < 0.1, assay.type="Cosine", batch=colData(joint.sce)$Batch, d=50,
                                                    preserve.single=TRUE)[[1]]

joint.meta <- data.frame("CellID"=colnames(joint.sce), "Batch"=colData(joint.sce)$Batch)
joint.sce <- runUMAP(joint.sce, n_neighbors=21, dimred="Cosine.PCA", n_dimred=30)
joint.umap <- data.frame("CellID"=colnames(joint.sce), "UMAP1"=reducedDim(joint.sce, "UMAP")[, 1],
                            "UMAP2"=reducedDim(joint.sce, "UMAP")[, 2])
joint.umap <- merge(joint.umap, joint.meta, by="CellID")
```

Now for the batch integration.

```{r "integrate Perinatal and Wells data"}
set.seed(42)
joint.mnn <- reducedMNN(reducedDim(joint.sce, "Cosine.PCA"), batch=colData(joint.sce)$Batch, k=21)
reducedDim(joint.sce, "Cosine.MNN") <- joint.mnn$corrected
joint.sce <- runUMAP(joint.sce, n_neighbors=21, dimred="Cosine.MNN", n_dimred=30, name="MNN.UMAP")

joint.mnn.umap <- data.frame("CellID"=colnames(joint.sce), "UMAP1"=reducedDim(joint.sce, "MNN.UMAP")[, 1],
                            "UMAP2"=reducedDim(joint.sce, "MNN.UMAP")[, 2])
joint.mnn.umap <- merge(joint.mnn.umap, joint.meta, by="CellID")
```

```{r, fig.height=6, fig.width=9}
pre.mnn <- ggplot(joint.umap, aes(x=UMAP1, y=UMAP2, colour=Batch)) +
  geom_scattermore(pointsize=1) +
  theme_cowplot() +
  scale_colour_npg() +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  NULL

post.mnn <- ggplot(joint.mnn.umap, aes(x=UMAP1, y=UMAP2, colour=Batch)) +
  geom_scattermore(pointsize=1) +
  theme_cowplot() +
  scale_colour_npg() +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  NULL

facet.mnn <- ggplot(joint.mnn.umap, aes(x=UMAP1, y=UMAP2, colour=Batch)) +
  geom_scattermore(data=joint.mnn.umap[, c("UMAP1", "UMAP2")],
                   pointsize=0.5, colour='grey80') +
  geom_scattermore(aes(colour=Batch), pointsize=1) +
  theme_cowplot() +
  scale_colour_npg() +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  facet_wrap(~Batch, nrow=1) +
  NULL

plot_grid(plot_grid(pre.mnn, post.mnn),
          facet.mnn, byrow=FALSE, ncol=1)

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/Wells_Perinatal_integration.png",
       height=6, width=9, dpi=300, bg='white')
```

The top left plot shows the data pre-integration, where as to the right
post-integration. The bottom panel is split by batch to show the
integration makes visual sense, i.e. the batches are not separated. The
next step is to use this MNN integrated space to perform a kNN label
transfer. We begin by building the joint kNN-graph and asking for each
cell, what the labels of the nearest neighbours are. For this to work we
would implicitly assume that there are mutual nearest neighbours for all
cells in both batches. We know this cannot be the case however, as the
Wells data are a subset of *mostly* mTEC-like cells. Therefore, I will
need to have most of the perinatal cells unlabelled.

```{r "kNN label transfer Perinatal -> Wells"}
wells.cells <- which(colData(joint.sce)$Batch %in% c("Wells"))
perinatal.cells <- which(colData(joint.sce)$Batch %in% c("Perinatal"))
joint.milo <- Milo(joint.sce)
joint.milo <- buildGraph(joint.milo, k=21, d=30, reduced.dim="Cosine.MNN")
joint.graph <- miloR::graph(joint.milo)

wells.label.list <- list()
for(x in seq_along(wells.cells)){
  x.idx <- wells.cells[x]
  x.nn <- unlist(neighborhood(joint.graph, nodes=x.idx))
  x.peri <- colnames(joint.sce)[intersect(x.nn, perinatal.cells)]
  if(length(x.peri) > 1){
    x.nn.labels <- table(cell.to.nhgroup[x.peri, ]$UniqueLabel)
    x.max.lab <- names(which(x.nn.labels == max(x.nn.labels)))
  } else{
    x.max.lab <- "Unknown"
  }
  
  if(length(x.max.lab) > 1){
    # randomly sample
    x.label <- sample(x.max.lab, size=1)
  } else{
    x.label <- x.max.lab
  }
  
  wells.label.list[[x]] <- data.frame("Idx"=x.idx, "CellID"=colnames(joint.sce)[x.idx], "UniqueLabel"=x.label)
}

wells.label.df <- do.call(rbind.data.frame, wells.label.list)
table(wells.label.df$UniqueLabel)
```

Only 1 unknown cell - that's pretty good. Now I'll cross-compare these
to the labels from Wells *et al.*.

```{r "compare original labels to kNN labels"}
wells.compare <- merge(wells.label.df, all.meta, by.x="CellID", by.y="X")
table(wells.compare$UniqueLabel, wells.compare$at_stage)
```


```{r "construct data frame for alluvial plots"}
wells.to.perinatal <- as.data.frame(table(wells.compare$UniqueLabel, wells.compare$at_stage))

colnames(wells.to.perinatal) <- c("Perinatal.Group", "Wells", "Freq")
wells.to.perinatal$Perinatal <- as.character(wells.to.perinatal$Perinatal.Group)
wells.to.perinatal$Perinatal <- gsub(unlist(lapply(strsplit(wells.to.perinatal$Perinatal, split=" ", fixed=TRUE),
                                                  FUN=function(PX) paste(tail(PX, 1)))),
                                    pattern="(\\()([0-9]+)(\\))", replacement="\\2")
wells.to.perinatal$Perinatal.Group <- factor(gsub(as.character(wells.to.perinatal$Perinatal.Group), pattern="Perinatal_", replacement=""),
                                             levels= c("Immature cTEC (1)", "Cycling cTEC (2)", "cTEC late (3)", "cTEC early (4)", "mTEC (5)", "Intertypical (6)",
                                                       "Post-Aire mTEC (7)", "cTEC early (8)", "ILC-like (9)", "Cycling mTEC (10)", "Sca1+ mTEC (11)", "cTEC late (12)",
                                                       "Intertypical cTEC-biased (13)", "ILC-like (14)", "Cycling Intertypical (15)", "Tuft-like mTEC (16)", "ILC-like (17)",
                                                           "Aire+ mTEC (18)"))
wells.to.perinatal$Ageing <- factor(gsub(wells.to.perinatal$Wells, pattern="(\\S+_)([0-9]+)", replacement="\\2"),
                                    levels=c("Late_Aire", "Aire_positive", "Early_Aire", "Ccl21a_high", "Cortico_medullary", "Tuft"))

max.peri <- length(unique(wells.to.perinatal$Perinatal))
wells.to.perinatal$Perinatal <- factor(wells.to.perinatal$Perinatal,
                                levels=c(1:max.peri))

# convert to lodes form to use geom_flow and ggrepel
alluvium.lodes <- to_lodes_form(wells.to.perinatal, axes=c(2, 4), key="Dataset", id="Nhood", value=Group)
```


```{r "alluvium plot of label transfer", fig.height=7.15, fig.width=9.95}
ggplot(alluvium.lodes, aes(x=Dataset, stratum=Group, alluvium=Nhood, y=Freq)) +
    geom_flow(aes(fill=Perinatal.Group), width=1/8) +
    geom_stratum(fill='grey', color = "black", width=1/12) +
    theme_cowplot() + 
    geom_text_repel(aes(label = ifelse(as.numeric(Dataset) == 2, as.character(Group), NA)),
                    stat = "stratum", size = 4, direction = "y", nudge_x = -0.1) +
    geom_text_repel(aes(label = ifelse(as.numeric(Dataset) == 1, as.character(Group), NA)),
                    stat = "stratum", size = 4, direction = "y", nudge_x = 0.1) +
    scale_x_discrete(limits = c("Perinatal", "Wells"), expand = c(0.1, 0)) +
    scale_fill_manual(values=rev(nhoodlabel.cols)) +
    guides(fill=guide_legend(title="Perinatal\nTEC Group", override.aes=list(alpha=1, colour='black'), ncol=1)) +
    theme(axis.line=element_blank(), axis.text.y=element_blank(),
          axis.title=element_blank(), axis.text.x=element_text(size=20), axis.ticks=element_blank()) +
    NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/Alluvium_WellstoPerinatal_NhoodGroupMap.png",
       height=7.15, width=9.95, dpi=300, bg='white')
```

Here we can see that almost all of the `Ccl21a_high` group correspond to
Intertypical and cTEC-biased Intertypical TEC. In their eLife manuscript
Wells *et al.* describe a `TAC-TEC` population, but none of these labels
appear here. By inspecting the UMAP and the adjacent populations it
looks like these are the 'Early_Aire' population in the meta-data. In
which case in our data \~50% of these cells correspond to Aire+ mTEC,
25% are Intertypical TEC and the other 25% are Cycling Intertypical TEC.

```{r "visualise labels on UMAP", fig.height=4.25, fig.width=9}
label.df <- do.call(rbind.data.frame, list(wells.compare[, c("CellID", "UniqueLabel")], as.data.frame(cell.to.nhgroup[, c("CellID", "UniqueLabel")])))
joint.meta.merge <- merge(joint.mnn.umap, label.df, by='CellID')


ggplot(joint.meta.merge, aes(x=UMAP1, y=UMAP2)) +
  geom_scattermore(data=joint.mnn.umap[, c("UMAP1", "UMAP2")],
                   pointsize=0.5, colour='grey90') +
  geom_scattermore(aes(colour=UniqueLabel), pointsize=1) +
  theme_cowplot() +
  scale_colour_manual(values=nhoodlabel.cols) +
  guides(colour=guide_legend(override.aes=list(size=2.5))) +
  facet_wrap(~Batch, nrow=1) +
  NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/Wells_Perinatal_integratedUMAP.png",
       height=4.5, width=9, bg='white', dpi=300)
```

```{r "visualise Wells labels on UMAP", fig.height=4.5, fig.width=6}
wells.meta.merge <- merge(joint.meta.merge, wells.compare[, c("CellID", "UniqueLabel", "stage")], by=c('CellID', 'UniqueLabel'))


ggplot(wells.meta.merge,
       aes(x=UMAP1, y=UMAP2)) +
  geom_scattermore(data=joint.mnn.umap[, c("UMAP1", "UMAP2")],
                   pointsize=0.5, colour='grey90') +
  geom_scattermore(aes(colour=stage), pointsize=1) +
  theme_cowplot() +
  scale_colour_d3() +
  guides(colour=guide_legend(title="Wells label",
                             override.aes=list(size=2.5))) +
  NULL

ggsave("~/Dropbox/AgeingExperiment/Perinatal_manuscript/figures/Wells_integratedUMAP.png",
       height=4.5, width=6, bg='white', dpi=300)
```
